---
title: "Dynatrace Analytics Workshop - datamodel"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r packages, include = FALSE}
library(tidyverse)
library(skimr)

# datamodel
library(datamodelr)
library(DiagrammeR)
```


```{r data import, cache = TRUE}
# import csvs
filenames = list.files(path = "Olist_data/", pattern="*.csv")

filepath <- paste0("Olist_data/", filenames)

data <- lapply(filepath, read.csv) #import data in list

#renaming
library(stringr)
df_names <- str_extract(filenames, '.*(?=\\.csv)')

names(data) <- df_names
```

```{r}
# summary of data frames
lapply(data, skim)

```

Findings of first exploration: 

+ Table 1: Customer information  of 96.096 unique IDs including geolocation information
+ Table 2: Geoinformation that provides lat/lng coordinates for zipcodes
+ Table 3: Purchase data of 98.666 orders on position level (product) including seller information and prices 
+ Table 4: Payment information of 99.440 order
+ Table 5: Review information on 99.441 orders including reveiw messages, scores and timestamps
+ Table 6: Purchase data of 99.441 orders on order level including customer and order status information
+ Table 7: Product information of 32.951 product IDs
+ Table 8: Seller information of 3.095 seller IDs
+ Taable 9: Product category informatio that maps to english translation

```{r}

# create datamodel
datamodel <- dm_from_data_frames(data) %>% 
  dm_add_references( # add keys
    ., 
    olist_orders_dataset$customer_id == olist_customers_dataset$customer_id,
    olist_orders_dataset$order_id == olist_order_reviews_dataset$order_id,
    olist_orders_dataset$order_id ==olist_order_payments_dataset$order_id,
    olist_orders_dataset$order_id == olist_order_items_dataset$order_id,
    olist_order_items_dataset$seller_id == olist_sellers_dataset$seller_id,
    olist_orders_items_dataset$product_id == olist_products_dataset$product_id,
    olist_sellers_dataset$seller_zip_code_prefix == olist_geolocation_dataset$geolocation_zip_code_prefix,
    olist_customers_dataset$customer_zip_code_prefix == olist_geolocation_dataset$geolocation_zip_code_prefix,
    olist_products_dataset$product_category_name == product_category_name_translation$`ï..product_category_name`

    )

# graph datamodel
dm_create_graph(datamodel, rankdir = "BT", col_attr = c("column", "type")) %>% 
  dm_render_graph()
```

