---
title: "Olist Marketing Funnel"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```


```{r packages, include = FALSE, message = FALSE}
library(tidyverse)
library(skimr)
library(stringr)
library(lubridate)
library(DT)
library(plotly)

source("R/functions.R")
```

# Data import

Import data in list via function Calll to import_csv_data().

```{r data import, cache = TRUE}
data <- import_csv_data(path = "Olist_data/")
data_marketing <- import_csv_data(path = "Olist_Marketing_data/")
```

# Overview of data

Source https://www.kaggle.com/jungjoonlee/eda-with-ecommerce-marketplace-seller-side


>Marketing Qualified Lead (MQL) means a potential reseller/manufacturer who has an interest in selling their products on Olist. After a MQL filled a form on landing page to sign up for seller, a Sales Development Representative(SDR) contacted the MQL and gathered more information about the lead. Then a Sales Representative(SR) consulted the MQL. So interaction between SDRs/SRs and MQLs can affect conversion from MQLs to sellers. A MQL who finally signed up for seller is called a closed deal.

## Closed Deals

```{r}
skim(data_marketing$olist_closed_deals_dataset)
head(data_marketing$olist_closed_deals_dataset) %>% datatable(., rownames = FALSE, options = list(scrollX = TRUE))
```

## MQL

```{r}
skim(data_marketing$olist_marketing_qualified_leads_dataset)
head(data_marketing$olist_marketing_qualified_leads_dataset) %>% datatable(., rownames = FALSE, options = list(scrollX = TRUE))
```

# EDA Closed Deals

## Data wrangling

```{r}
# merge dataframes
df_deals <- data_marketing$olist_closed_deals_dataset %>% 
  right_join(data_marketing$olist_marketing_qualified_leads_dataset, by = "mql_id")

# createa monthly dataframe for deals by origin by first contact date
df_deals_monthly <- df_deals %>% 
  mutate(first_contact_month = format.Date(as_date(first_contact_date), "%Y%m"),
         deal_closed = if_else(is.na(seller_id), 0, 1),
         origin = if_else(origin == "", "unknown", as.character(origin))) %>% 
  group_by(origin, first_contact_month) %>% 
  summarize(leads = n(),
            closed_deals = sum(deal_closed)) %>% 
  ungroup()
  
```

# Analysis Leads

## Leads Overall

```{r}

plot_leads <- df_deals_monthly %>% 
  group_by(first_contact_month) %>% 
  summarize(leads = sum(leads)) %>% 
  ggplot(., aes(x = first_contact_month, y = leads, group = 1)) +
  geom_line() +
  geom_point() +
  theme_bw() +
  labs(x = "", y = "Number of Leads", colour = "origin")

```

```{r, fig.width=11}
ggplotly(plot_leads)

#plot_leads
```


## Leads by Origin

```{r}

plot_leads_origin <- df_deals_monthly %>% 
  select(-closed_deals) %>% 
  pivot_wider(., names_from = origin, values_from = leads) %>% 
  replace(is.na(.), 0) %>% 
  ggplot(., aes(x = first_contact_month, group = 1)) +
  geom_line(aes(y = direct_traffic, color = "direct_traffic")) +
  geom_line(aes(y = display, color = "display"))+
  geom_line(aes(y = email, color = "email")) +
  geom_line(aes(y = organic_search, color = "organic_search")) +
  geom_line(aes(y = paid_search, color = "paid_search")) +
  geom_line(aes(y = referral, color = "referral")) +
  geom_line(aes(y = social, color = "social")) +
  geom_line(aes(y = other_publicities, color = "other_publicities")) +
  geom_line(aes(y = other, color = "other")) +
  geom_line(aes(y = unknown, color = "unknown")) +
  geom_point(aes(y = direct_traffic, color = "direct_traffic")) +
  geom_point(aes(y = display, color = "display"))+
  geom_point(aes(y = email, color = "email")) +
  geom_point(aes(y = organic_search, color = "organic_search")) +
  geom_point(aes(y = paid_search, color = "paid_search")) +
  geom_point(aes(y = referral, color = "referral")) +
  geom_point(aes(y = social, color = "social")) +
  geom_point(aes(y = other_publicities, color = "other_publicities")) +
  geom_point(aes(y = other, color = "other")) +
  geom_point(aes(y = unknown, color = "unknown")) +
  theme_bw() +
  labs(x = "", y = "Number of Leads", colour = "origin")
```

```{r, fig.width=11}
ggplotly(plot_leads_origin)

#plot_leads_origin
```

```{r}
plot_bar_leads_origin <- df_deals_monthly %>% 
  select(-closed_deals) %>% 
  group_by(origin) %>% 
  summarize(leads = sum(leads)) %>% 
  ggplot(., aes(x = reorder(origin, -leads), y = leads, group = 1)) +
  geom_col(aes(fill = origin))+
  geom_text(aes(label = ..y..), vjust = -1) +
  theme_bw() +
  labs(x = "", y = "Number of Leads", colour = "origin") +
  scale_y_continuous(limits = c(0, 2350))
```

```{r, fig.width=11}
ggplotly(plot_bar_leads_origin)

#plot_bar_leads_origin
```

```{r}
plot_bar_leads_origin_per <- df_deals_monthly %>% 
  select(-closed_deals) %>% 
  group_by(origin) %>% 
  summarize(leads = sum(leads)) %>% 
  ggplot(., aes(x = reorder(origin, -leads), y = leads / sum(leads), group = 1)) +
  geom_col(aes(fill = origin))+
  geom_text(aes(label = paste0(round(..y.. * 100,1), "%")), vjust = -1) +
  theme_bw() +
  labs(x = "", y = "Number of Leads", colour = "origin") +
  scale_y_continuous(labels=scales::percent, limits = c(0, 0.3)) 
```

```{r, fig.width=11}
ggplotly(plot_bar_leads_origin_per)

#plot_bar_leads_origin_per
```

# Analysis Closed Deals

## Closed Deals Overall

```{r}

plot_leads <- df_deals_monthly %>% 
  group_by(first_contact_month) %>% 
  summarize(leads = sum(closed_deals)) %>% 
  ggplot(., aes(x = first_contact_month, y = leads, group = 1)) +
  geom_line() +
  geom_point() +
  theme_bw() +
  labs(x = "", y = "Number of Closed Deals", colour = "origin")

```

```{r, fig.width=11}
ggplotly(plot_leads)

# plot_leads
```


## Leads by Origin

```{r}

plot_leads_origin <- df_deals_monthly %>% 
  select(-leads) %>% 
  pivot_wider(., names_from = origin, values_from = closed_deals) %>% 
  replace(is.na(.), 0) %>% 
  ggplot(., aes(x = first_contact_month, group = 1)) +
  geom_line(aes(y = direct_traffic, color = "direct_traffic")) +
  geom_line(aes(y = display, color = "display"))+
  geom_line(aes(y = email, color = "email")) +
  geom_line(aes(y = organic_search, color = "organic_search")) +
  geom_line(aes(y = paid_search, color = "paid_search")) +
  geom_line(aes(y = referral, color = "referral")) +
  geom_line(aes(y = social, color = "social")) +
  geom_line(aes(y = other_publicities, color = "other_publicities")) +
  geom_line(aes(y = other, color = "other")) +
  geom_line(aes(y = unknown, color = "unknown")) +
  geom_point(aes(y = direct_traffic, color = "direct_traffic")) +
  geom_point(aes(y = display, color = "display"))+
  geom_point(aes(y = email, color = "email")) +
  geom_point(aes(y = organic_search, color = "organic_search")) +
  geom_point(aes(y = paid_search, color = "paid_search")) +
  geom_point(aes(y = referral, color = "referral")) +
  geom_point(aes(y = social, color = "social")) +
  geom_point(aes(y = other_publicities, color = "other_publicities")) +
  geom_point(aes(y = other, color = "other")) +
  geom_point(aes(y = unknown, color = "unknown")) +
  theme_bw() +
  labs(x = "", y = "Number of Closed Deals", colour = "origin")
```

```{r, fig.width=11}
ggplotly(plot_leads_origin)

# plot_leads_origin
```

```{r}
plot_bar_leads_origin <- df_deals_monthly %>% 
  select(-leads) %>% 
  group_by(origin) %>% 
  summarize(leads = sum(closed_deals)) %>% 
  ggplot(., aes(x = reorder(origin, -leads), y = leads, group = 1)) +
  geom_col(aes(fill = origin))+
  geom_text(aes(label = ..y..), vjust = -1) +
  theme_bw() +
  labs(x = "", y = "Number of Closed Deals", colour = "origin") +
  scale_y_continuous(limits = c(0, 300))
```

```{r, fig.width=11}
ggplotly(plot_bar_leads_origin)

# plot_bar_leads_origin
```

```{r}
plot_bar_leads_origin_per <- df_deals_monthly %>% 
  select(-leads) %>% 
  group_by(origin) %>% 
  summarize(closed_deals = sum(closed_deals)) %>% 
  ggplot(., aes(x = reorder(origin, -closed_deals), y = closed_deals / sum(closed_deals), group = 1)) +
  geom_col(aes(fill = origin))+
  geom_text(aes(label = paste0(round(..y.. * 100,1), "%")), vjust = -1) +
  theme_bw() +
  labs(x = "", y = "Number of Closed Deals", colour = "origin") +
  scale_y_continuous(labels=scales::percent, limits = c(0, 0.35)) 
```

```{r, fig.width=11}
ggplotly(plot_bar_leads_origin_per)

# plot_bar_leads_origin_per
```

# Conversion Rate

## Closed Deals Overall

```{r}

plot_conversion <- df_deals_monthly %>% 
  group_by(first_contact_month) %>% 
  summarize(leads = sum(leads),
            closed_deals = sum(closed_deals)) %>% 
  mutate(conversion_rate = closed_deals / leads) %>% 
  ggplot(., aes(x = first_contact_month, y = conversion_rate, group = 1)) +
  geom_line() +
  geom_point() +
  theme_bw() +
  labs(x = "", y = "Conversion Rate", colour = "origin") +
  scale_y_continuous(labels=scales::percent) 

```

```{r, fig.width=11}
ggplotly(plot_conversion)

# plot_conversion
```

## Conversion Rate by Origin

### 2017

```{r}
plot_bar_conversion_origin_2017 <- df_deals_monthly %>% 
  filter(substr(first_contact_month,1,4) == "2017") %>% 
  group_by(origin) %>% 
  summarize(leads = sum(leads),
            closed_deals = sum(closed_deals)) %>% 
  mutate(conversion_rate = closed_deals / leads) %>% 
  ggplot(., aes(x = reorder(origin, -conversion_rate), y = conversion_rate, group = 1)) +
  geom_col(aes(fill = origin))+
  geom_text(aes(label = paste0(round(..y.. * 100,1), "%")), vjust = -1) +
  theme_bw() +
  labs(x = "", y = "Conversion Rate", colour = "origin") +
  scale_y_continuous(labels=scales::percent, limits = c(0, 0.20))
```

```{r, fig.width=11, fig.height=3}
ggplotly(plot_bar_conversion_origin_2017)

# plot_bar_conversion_origin_2017
```

### 2018

```{r}
plot_bar_conversion_origin_2018 <- df_deals_monthly %>% 
  filter(substr(first_contact_month,1,4) == "2018") %>% 
  group_by(origin) %>% 
  summarize(leads = sum(leads),
            closed_deals = sum(closed_deals)) %>% 
  mutate(conversion_rate = closed_deals / leads) %>% 
  ggplot(., aes(x = reorder(origin, -conversion_rate), y = conversion_rate, group = 1)) +
  geom_col(aes(fill = origin))+
  geom_text(aes(label = paste0(round(..y.. * 100,1), "%")), vjust = -1) +
  theme_bw() +
  labs(x = "", y = "Conversion Rate", colour = "origin") +
  scale_y_continuous(labels=scales::percent, limits = c(0, 0.30))
```

```{r, fig.width=11, fig.height=3}
ggplotly(plot_bar_conversion_origin_2018)

# plot_bar_conversion_origin_2018
```
